---
title: "DATA 608: Knowledge And Visual Analytics"
author: "Anjal Hussan, Fall 2020"
subtitle: "HW1 - ggplot2"
output:
  html_document:
    highlight: zenburn
    theme: lumen
    toc: true
    toc_float: true
---

**Principles of Data Visualization and Introduction to ggplot2**

# Environment setup

```{r, message=F, warning=F}
if (!require('dplyr')) install.packages('dplyr')
if (!require('ggplot2')) install.packages('ggplot2')
if (!require('scales')) install.packages('scales')
```

# Data loading

Read the data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

myDataSet <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}
head(myDataSet)
```
```{r}

summary(myDataSet)

```


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

### Answer

```{r}

groupByStateWise <- myDataSet %>% 
    group_by(State) %>% 
    tally()
ggplot(groupByStateWise, aes(x=reorder(State, n), y=n)) + 
    geom_point(size=1) + 
    geom_segment(aes(x=State, xend=State, y=0, yend=n)) +
    labs(title="Distribution of 5000 Fastest Growing Companies by State", 
         caption="Source: Inc Magazine",
         y="Counts", 
         x="State") + 
    theme_classic(base_size = 8) +
    coord_flip()
ggsave('FastestGrowingCompanyByState.png')

```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

### Answer

```{r}

# Answer Question 2 here
third_state <- subset(myDataSet, State=="NY")
third_state <- third_state[complete.cases(third_state$Industry), ] 
third_state <- third_state[complete.cases(third_state$Employees), ] 

third_state <- arrange(third_state, desc(Employees))
lower <- min(third_state$Employees)
upper <- max(third_state$Employees)
meidan_ny <- paste0("Median NY Employees:  ", median(third_state$Employees))
ggplot(third_state, aes(x=reorder(Industry, Employees, FUN=median), y=Employees)) + 
    geom_boxplot(outlier.shape = NA) + 
    scale_y_continuous(trans = log2_trans(), limits = c(lower, upper)) +
    geom_hline(yintercept = median(third_state$Employees),
               color="grey", 
               linetype="dashed") +
    geom_text(aes(x=1.5, label=meidan_ny, y = 280), 
              size = 3,
              colour="grey") +
    labs(title="Number of Employees By Industry in NY State",
         subtitle="5000 Fastest Growing Companies by Industry",
         caption="Source: Inc Magazine",
         y="Num Employees, Log2 Transform", 
         x="") + 
    theme_classic() +
    coord_flip()
ggsave('third_state_Median_employment.png')

```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

### Answer
```{r}

revenue <- myDataSet[complete.cases(myDataSet$Revenue), ]
revenue <- revenue[complete.cases(revenue$Employees), ] 
revenue <- revenue %>%
    group_by(Industry) %>%
    summarise(TotRev = sum(Revenue, na.rm=T), TotEmp = sum(Employees, na.rm=T)) %>%
    mutate(AvgRevK = (TotRev/TotEmp)/1000)
ggplot(revenue, aes(x=reorder(Industry, AvgRevK), y=AvgRevK)) + 
    geom_bar(stat="identity", width=.5, fill="tomato3")+ 
    labs(title="Revenue Per Employee",
         subtitle="5000 Fastest Growing Companies by Industry", 
         caption="Source: Inc Magazine",
         y="Revenue - Thousands ($)", 
         x="") + 
    theme_classic() +
    coord_flip()
ggsave('revenue_per_employee.png')
```

### Help:
* http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html
* http://www.cookbook-r.com/Graphs/Axes_(ggplot2)/
